A quantum parallel Markov chain Monte Carlo

### 2021-11-28 main.tex
intro done

\emph{In the presence of these trade-offs, we seek to enjoy the benefits of parallel MCMC while (\textcolor{red}{a}) limiting the total amount of computing necessary and (\textcolor{red}{b})  retaining parallel computing resources for computational bottlenecks such as proposal generations and density evaluations. Here, we assert that established quantum algorithms can help achieve these twin goals.}

Quantum computers 
with gates that are mathematically equivalent to unitary operators \citep{nielsen2002quantum}.  Assuming that engineers overcome the tremendous difficulties involved in building a practical quantum computer (where practicality entails simultaneous use of many quantum gates with little additional noise), 21st century statisticians might have access to quadratic or even exponential speedups for extremely specific statistical tasks.  We are particularly interested in the following four quantum algorithms: quantum search \citep{grover1996fast}, or finding a single 1 amid a collection of 0s, only requires $\order{\sqrt{N}}$ queries, delivering a quadratic speedup over classical search; quantum counting \citep{boyer1998tight}, or finding the number of 1s amid a collection of 0s, only requires $\order{\sqrt{N/M}}$ (where $M$ is the number of 1s) and could be useful for generating p-values within Monte Carlo simulation from a null distribution; to obtain the gradient of a function (e.g., the log-likelihood for Fisher scoring or HMC) with a quantum computer, one only needs to evaluate the function once \citep{jordan2005fast} as opposed to $\order(P)$ times for numerical differentiation, and there is nothing stopping the statistician from using, say, a GPU for this single function call; finally, the HHL algorithm \citep{harrow2009quantum} obtains the scalar value

\section{Bare-bones quantum computing}

@@

\section{Quantum search and quantum minimization}

@@

\section{Parallel MCMC}

@@

\subsection{Gumbel-max}
\newcommand{\p}{\mbox{p}}

@@

\subsection{qp$^2$MCMC}

>>>>

\emph{In the presence of these trade-offs, we seek to enjoy the benefits of parallel MCMC while (\textcolor{red}{a}) limiting the total amount of computing necessary and (\textcolor{red}{b})  retaining conventional parallel computing resources for secondary computational bottlenecks such as proposal generations and proposal density evaluations. Here, we assert that established quantum algorithms can help achieve these twin goals.}

Quantum computers use quantum mechanics to store information and manipulate data. By leveraging the idiosyncracies of quantum physics, these computers are able to deliver speedups over conventional computing for a relatively small number of computational problems. Some of these speedups are very large. Quantum computers can achieve logarithmic complexity for some classically linear time algorithms, turning $\order{N}$ into $\order{\log N}$.  Shor's quantum algorithm for integer factorization \citep{shor} and the HHL algorithm for linear system solving \citet{harrow2009quantum} deliver just such an \emph{exponential} speedup \citep{shor}. Other algorithms deliver a still impressive \emph{polynomial} speedup.  For example, the algorithms considered in the following (Section \ref{sec:qmin}) achieve quadratic speedups over conventional computing, turning  $\order{N}$ runtimes into $\mathcal{O}(\sqrt{ N})$.  
 Both the magnitude of quantum computing's successes and the relative rarity of those successes mean that there is a general interest in leveraging quantum algorithms for previously unconsidered computational problems.  We will show that---with the help of the Gumbel-max trick (Section \ref{sec:gm})---established quantum optimization techniques are actually useful for sampling from computationally expensive discrete distributions and apply this insight to achieving quadratic speedups for parallel MCMC.

In the following section, we provide a \emph{very} short introduction to a few of the basic elements of quantum computing. The interested reader may look to \citet{nielsen2002quantum} for a canonical introduction to quantum algorithms and their implementation.

\section{Bare-bones quantum computing}\label{sec:shortIntro}

@@

\section{Quantum search and quantum minimization}\label{sec:qmin}

@@

\section{Quantum parallel MCMC}

@@

\subsection{Gumbel-max}\label{sec:gm}
\newcommand{\p}{\mbox{p}}

@@

\subsection{QPMCMC}
